Area maximisation in continuous time

### 2013-11-07 Notes/ContTime.tex
typos
Laplace transform of SNR

We can rewrite this using \eqref{eq:diffpT}, with $A=\sqrt{N}(2f^+f^-)\hat{A}$ and $\enc_{ij} \equiv \diff{\frg_{ij}}{\alpha}=\W^+_{ij}-\W^-_{ij}$
%
\begin{equation}\label{eq:areaT}
  \hat{A} = \sum_{i,j,k} \eq_i \enc_{ij} (\fptb_{ik} - \fptb_{jk}) \eq_k \w_k.
\end{equation}
%

@@

The SNR at time $t$ was given in \eqref{eq:SNRcurve} as
%
\begin{equation*}
  \SNR(t) = \sqrt{N} (2f^+f^-) \eq \enc \, \e^{rt\frg}\w.
\end{equation*}
%
Let us try to maximise this at time $t$ only.
We will consider the independent parameters to be the off-diagonal elements of $\M^\pm$, with the diagonal elements determined via $\M^\pm_{ii} = \sum_{j\neq i} \M^\pm_{ij}$.
Consider the Lagrangian
%
\begin{equation}\label{eq:finlagrangian}
  \CL = \eq \enc\e^{rt\frg}\w + \sum_\pm \sum_{ij} \mu^\pm_{ij} \M^\pm_{ij}.
\end{equation}
%
Necessary conditions for an extremum are
%
\begin{equation*}
  \pdiff{\CL}{\M^\pm_{gh}} = 0,
  \qquad
    \mu^\pm_{gh} \geq 0,\quad
    \M^\pm_{gh} \geq 0,\quad
    \mu^\pm_{gh}\M^\pm_{gh} = 0.
\end{equation*}
%
We can use the results of \cite{Ortiz2001diffmatexp} to write down the shift of this Lagrangian under an infinitesimal perturbation
%
\begin{equation}\label{eq:finpert}
\begin{aligned}
  \delta\CL =&\, \eq \delta\frg \fund\, \enc \, \e^{rt\frg} \w
    + \eq \delta\enc \, \e^{rt\frg} \w %\\ &
   + r\int_{0}^{t}\!\!\dr\tau\, \eq \enc\, \e^{r\tau\frg} \delta\frg \e^{r(t-\tau)\frg} \w\\
   &+ \sum_\pm \sum_{ij} \mu^\pm_{ij} \delta\M^\pm_{ij}
\end{aligned}
\end{equation}
%
This can be rewritten using the eigenvector decomposition of $\frg$.
%
\begin{equation}\label{eq:eigendecomp}
  \frg = \sum_a -q_a \uv^a \vv^a,
  \quad
  \vv^a \uv^b = \delta_{ab},
  \quad
  \frg \uv^a = -q_a \uv^a,
  \quad
  \vv^a \frg = -q_a \vv^a.
\end{equation}
%
People often write $\vv$ as $\vv\trans$, presumably because they find the concept of row vectors too difficult to handle.
We're better than that.
We can choose $q_1=0$, $\uv^1=\onev$ and $\vv=\eq$.

@@

@@

This means that the derivatives of the area are:
%
\begin{equation}\label{derivTarea}
  \pdiff{\hat{A}}{\fptb_{gh}} = -\eq_h\brk{
    \sum_i \eq_i \enc_{ig} \w_h
    + r\sum_{ijk} \prn{\frg_{ig}+\frg_{kg}} \eq_i \enc_{ij} \prn{\fptb_{ik}-\fptb_{jk}} \eq_k \w_k
    }.
\end{equation}
%

>>>>

We can rewrite this using \eqref{eq:diffpT}, with $rA=\sqrt{N}(2f^+f^-)\hat{A}$ and $\enc_{ij} \equiv \diff{\frg_{ij}}{\alpha}=\W^+_{ij}-\W^-_{ij}$
%
\begin{equation}\label{eq:areaT}
  \hat{A} = \sum_{i,j,k} \eq_i \enc_{ij} (\fptb_{ik} - \fptb_{jk}) \eq_k \w_k.
\end{equation}
%

@@

The SNR at time $t$ was given in \eqref{eq:SNRcurve} as
%
\begin{equation*}
  \SNR(t) = \sqrt{N} (2f^+f^-) \eq \enc \, \e^{rt\frg}\w.
\end{equation*}
%
Let us try to maximise this at time $t$ only.
We will consider the independent parameters to be the off-diagonal elements of $\M^\pm$, with the diagonal elements determined via $\M^\pm_{ii} = \sum_{j\neq i} \M^\pm_{ij}$.
Consider the Lagrangian
%
\begin{equation}\label{eq:finlagrangian}
  \CL = \eq \enc\e^{rt\frg}\w + \sum_\pm \sum_{ij} \mu^\pm_{ij} \M^\pm_{ij}.
\end{equation}
%
Necessary conditions for an extremum are
%
\begin{equation*}
  \pdiff{\CL}{\M^\pm_{gh}} = 0,
  \qquad
    \mu^\pm_{gh} \geq 0,\quad
    \M^\pm_{gh} \geq 0,\quad
    \mu^\pm_{gh}\M^\pm_{gh} = 0.
\end{equation*}
%
We can use the results of \cite{Ortiz2001diffmatexp} to write down the shift of this Lagrangian under an infinitesimal perturbation
%
\begin{equation}\label{eq:finpert}
\begin{aligned}
  \delta\CL =&\, \eq \delta\frg \fund\, \enc \, \e^{rt\frg} \w
    + \eq \delta\enc \, \e^{rt\frg} \w %\\ &
   + r\int_{0}^{t}\!\!\dr\tau\, \eq \enc\, \e^{r\tau\frg} \delta\frg \e^{r(t-\tau)\frg} \w\\
   &+ \sum_\pm \sum_{ij} \mu^\pm_{ij} \delta\M^\pm_{ij}
\end{aligned}
\end{equation}
%
This can be rewritten using the eigenvector decomposition of $\frg$.
%
\begin{equation}\label{eq:eigendecomp}
  \frg = \sum_a -q_a \uv^a \vv^a,
  \quad
  \vv^a \uv^b = \delta_{ab},
  \quad
  \frg \uv^a = -q_a \uv^a,
  \quad
  \vv^a \frg = -q_a \vv^a.
\end{equation}
%
People often write $\vv$ as $\vv\trans$, presumably because they find the concept of row vectors too difficult to handle.
We're better than that.
We can choose $q_1=0$, $\uv^1=\onev$ and $\vv^1=\eq$.

@@

\section{Laplace transform}\label{sec:laplace}

\subsection{Fundamental matrix \etc}\label{sec:lfund}

In analogy to \eqref{eq:funddef}, define
%
\begin{equation}\label{eq:lfund}
  \fund(s) = (s\I +\onev\pib -\MM)\inv.
\end{equation}
%
This reduces to the fundamental matrix at $s=0$.
The analogous result to \eqref{eq:fundrowsum} is
%
\begin{equation}\label{eq:lfundrowsum}
  \fund(s)\onev = \frac{\tau}{1+s\tau}\onev.
\end{equation}
%
The analogues of \eqref{eq:fundprob}, \eqref{eq:fundQZ} and \eqref{eq:fundZQ} are much uglier, but luckily we won't need them.
Supose we have some row vector $\mathbf{a}$ such that $\mathbf{a}\onev=0$.
Then $\mathbf{a}\fund(s)$ is independent of $\pib$:
%
\begin{equation}\label{eq:alfund}
  \pdiff{(\mathbf{a}\fund(s))_i}{\pib_k} = (\mathbf{a}\fund(s)\onev) \fund_{ki}(s) = \frac{\tau}{1+s\tau} (\mathbf{a}\onev) \fund_{ki}(s) = 0.
\end{equation}
%

Then we also define
%
\begin{equation}\label{eq:lfptb}
  \fptb(s) = (\onem\fund\dg(s)-\fund(s))\D,
  \qquad
  \fptb_{ij}(s) = \frac{\fund_{jj}(s)-\fund_{ij}(s)}{\eq_j}.
\end{equation}
%
This reduces to the first passage time matrix at $s=0$, but at other $s$ there is (probably) no relation.

Using the exact same proof as \autoref{th:kemenyconst}, we see that the quantity
%
\begin{equation}\label{eq:lkemeny}
  \eta(s) = \sum_j \fptb_{ij}(s)\eq_j
\end{equation}
%
is independent of the starting state $i$, but now $\eta(s) = \tr\fund(s) - \frac{\tau}{1+s\tau}$.
We can the define 
%
\begin{equation}\label{eq:lpmixtime}
  \eta^{\pm}_i(s) = \sum_j \fptb_{ij}(s)\eq_j \prn{\frac{1\pm\w_j}{2}} =  \sum_{j\in\pm} \fptb_{ij}(s)\eq_j.
\end{equation}
%
We can arrange the states in order of decreasing $\eta^{+}_i(s)$ or increasing $\eta^{-}_i(s)$.

\subsection{Laplace transform of SNR curve}\label{sec:laplaceSNR}

Consider the Laplace transform of the evolution operator:
%
\begin{equation}\label{eq:levol}
  \mathbf{G}(s) = \int_0^{\infty}\dr t\, \e^{(\MM-s) t}.
\end{equation}
%
For $\Re s>0$, we have 
%
\begin{equation}\label{eq:levolcalc}
  (s-\MM)\mathbf{G}(s) = \int_0^{\infty}\dr t\, (s-\MM)\e^{(\MM-s) t} = \brk{-\e^{(\MM-s) t}}_0^{\infty} = \I.
\end{equation}
%
As $(s-\MM)$ is invertible for $\Re s>0$, because all eigenvalues of $\MM$ are nonpositive, we have 
%
\begin{equation}\label{eq:levolres}
  \mathbf{G}(s) = (s-\MM)\inv.
\end{equation}
%
For $s=0$, we can avoid our problems by replacing $\mathbf{G}(s) \to \fund(s)$.

Now consider the Laplace transform of the SNR curve \eqref{eq:SNRcurve}
%
\begin{equation}\label{eq:lsnr}
\begin{aligned}
  A(s) &= \int_0^{\infty}\dr t \e^{-st}\SNR(t) \\
   &= \int_0^{\infty}\dr t\, \sqrt{N}(2f^+f^-) \eq \enc \e^{(r\frg-s)t} \w \\
   &= \frac{\sqrt{N}(2f^+f^-)}{r} \eq \enc \prn{s - r\frg}\inv \w.
\end{aligned}
\end{equation}
%
Note that $A(0)=A$, the area, and $\lim_{s\to\infty} \brc{sA(s)} = \SNR(0)$, the initial SNR.

Thanks to \eqref{eq:alfund}, we can replace $\mathbf{G}(s) \to \fund(s)$, as $\enc\onev=0$.
%
\begin{equation}\label{eq:larea}
\begin{aligned}
  \hat{A}(s) = \frac{rA(s)}{\sqrt{N}(2f^+f^-)} &=  \eq \enc \fund(s) \w \\
    &= \sum_{ijk} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \eq_k \w_k \\
    &= 2\sum_{ij}  \eq_i \enc_{ij} (\eta^+_{i}(s) - \eta^+_{j}(s)).
\end{aligned}
\end{equation}
%
In analogy with \eqref{eq:areacoeffs}, we define
%
\begin{equation}\label{eq:lareacoeffs}
  \begin{aligned}
    c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \\
    a_i(s) &= \sum_j \eq_i \enc_{ij} (\eta^+_{i}(s) - \eta^+_{j}(s)),\\
    \implies
    \hat{A}(s) &= \sum_k c_k(s) \eq_k \w_k
      = 2\sum_i a_i(s).
  \end{aligned}
\end{equation}
%

\subsection{Perturbation analysis}\label{sec:lpert}

Using the same approach as the area bound, we find that the derivatives are
%
\begin{equation}\label{eq:lAdiff}
  \pdiff{\hat{A}(s)}{\W^\pm_{gh}} = 2rf^\pm \eq_g \brk{ \sum_i a_i(s) (\fptb_{gi}(0)-\fptb_{hi}(0))
     +c_g(s) (\eta^+_g(s) - \eta^+_h(s)) } \pm 2\eq_g (\eta^+_g(s) - \eta^+_h(s)).
\end{equation}
%
This is exactly the same as \eqref{eq:derivA}.
It follows that, using exactly the same perturbations, that the Laplace transform of the SNR curve for \emph{any} model is lower than that of some model with the serial topology having the same equilibrium distribution.

@@

This means that the derivatives of the area are:
%
\begin{equation}\label{eq:derivTarea}
  \pdiff{\hat{A}}{\fptb_{gh}} = -\eq_h\brk{
    \sum_i \eq_i \enc_{ig} \w_h
    + r\sum_{ijk} \prn{\frg_{ig}+\frg_{kg}} \eq_i \enc_{ij} \prn{\fptb_{ik}-\fptb_{jk}} \eq_k \w_k
    }.
\end{equation}
%
